---
title: "CTD binning"
author: "Marko Lipka"
date: "8/20/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("CTDbinning.R")
```

Function is defined in **CTDbinning.R**.

Function *binCTD.mean()* applies the *mean()* function to all parameters of a given data frame binned by one of its parameters (usually depth or pressure).

Arguments:

- **.data**: the data.frame to be summarized
- **.binvar**: the variable name to bin over
- **.breaks**: breaks as used by *cut()* (either a single value giving the number of equally distributed bins or a vector of cut points)
- **.binwidth**: alternatively to **.breaks** the binwidth can be set (overwrites **.breaks**)

# Example:

## Raw example data

... from a CTD cast in the Arkona Basin (southern Baltic Sea).
(... extracted from SBE CTD *.cnv* file with the helper function [**read.cnv.file()**](https://github.com/markolipka/readCNV))

```{r example}
load("example_ctd") # load dumped example data frame
ctd <- select(ctd, prDM : par) # reduce number of parameters for greater clarity

knitr::kable(head(ctd), digits = 1) # print table head

knitr::kable(tail(ctd), digits = 1) # print table tail
```

## Binned: 

### Example 1: binned by **.binwidth**
```{r binned_binwidth}
binned.binwidth <- binCTD.mean(.data = ctd,
                               .binvar = "prDM",
                               .binwidth = 1)

knitr::kable(binned.binwidth, digits = 1)
```

```{r plot_binwidth}
binned.binwidth %>%
    gather(key = "variable", value = "value", -bins, -prDM) %>%
    ggplot() +
    aes(x = prDM, y = value) +
    geom_vline(xintercept = 1:18, color = "deeppink") +
    geom_point(data = gather(ctd, key = "variable", value = "value", -prDM),
               shape = 21) +
    geom_point(color = "deeppink") +
    facet_grid(~variable, scales = "free", switch = "x") +
    coord_flip() +
    scale_x_reverse() +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = NA, colour = NA),
          strip.placement = "outside") +
    ylab(NULL) +
    ggtitle("Plot (binned vs. unbinned)")
```

### Example 2: binned by **.breaks**
```{r binned_breaks}
breaks <- c(0, 1:5, 10, 15, +Inf)
binned.breaks <- binCTD.mean(.data = ctd,
                               .binvar = "prDM",
                               .breaks = breaks)

knitr::kable(binned.breaks, digits = 1) # print table
```


```{r plot_breaks}
binned.breaks %>%
    gather(key = "variable", value = "value", -bins, -prDM) %>%
    ggplot() +
    aes(x = prDM, y = value) +
    geom_vline(xintercept = breaks, color = "deeppink") +
    geom_point(data = gather(ctd, key = "variable", value = "value", -prDM),
               shape = 21) +
    geom_point(color = "deeppink") +
    facet_grid(~variable, scales = "free", switch = "x") +
    coord_flip() +
    scale_x_reverse() +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = NA, colour = NA),
          strip.placement = "outside") +
    ylab(NULL) +
    ggtitle("Plot (binned vs. unbinned)")
```

# ToDo:

- might be useful to allow other functions than *mean()* for summarizing?!
